---
title: "BDA - Assignment 1"
author: "Anonymous"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 1
urlcolor: blue
---

```{r setup, include=FALSE}
# This chunk sets echo = TRUE as default, that is print all code.
# knitr::opts_chunk$set can be used to set other notebook generation options, too.
# include=FALSE inside curly brackets makes this block not be included in the pdf.
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# To install aaltobda, see the General information in the assignment.
remotes::install_github("avehtari/BDA_course_Aalto", subdir = "rpackage", upgrade = "never")
```

# Exercise 1 (Basic probability theory notation and terms): 

- **Probability:** Based on *known* parameters, how likely an specific event occurs.

- **Probability mass:** For discrete variables, how likely a series of events (could be a single event) occurs.

- **Probability density** For continuous variables, how likely an interval of events (probability at a single point is zero) occurs.

- **Probability mass function (pmf):** Related to *Probability mass*, it is the function that maps the event to the probability that it occurs.

- **Probability density function (pdf):** Related to *Probability density*, it is the function that maps the interval of events to the probability that it occurs.

- **Probability distribution:** The mass/density assigned by the pmf/pdf to each outcome occurrence.

- **Discrete probability distribution:** The mass assigned by the pmf to each outcome occurrence.  

- **Continuous probability distribution:** The density assigned by the pdf to each outcome occurrence.

- **Cumulative distribution function:** The summed of the mass/density from all outcomes occurrence as they are arranged from left to right.

- **Likelihood:** The probability of observing the data given the parameters. 

# Exercise 2 (Basic computer skills):

- a) 

```{r}
p  <- seq(0, 1, length = 100)
mu <- 0.2
sd <- .01
ps <-  dbeta(p, mu, sd)
plot(p, ps, type = "l")
```
- b) 
```{r}
n <- 1000
x <- rbeta(n, mu, sd)
hist(x)
```

- c)

```{r}
cat("This is the mean:", mean(x))
cat("\n")
cat("This is the variance:", var(x))
```

- d)

```{r}
quantile(x, prob = seq(0.1, 1, by = .05), CI.type = "two.sided")
```

# Exercise 3 (Bayes' theorem):

According from my notes for the exercise required in the pre-requisites, I found that $P(cancer does not get detected) = P(cancer | negative result) = 0.000002$, and $P(positive test but no cancer) = P(~cancer | positive result) = 0.97$. My conclusion was that while $P(C|N)$ is low, I am concerned about $P(~C|P)$ being close to 1. I would pass on investing on this company.

# Exercise 4 (Bayes' theorem):

```{r}
boxes <- matrix(c(2,4,1,5,1,3), 
                ncol = 2,
                dimnames = list(c("A", "B", "C"),
                                c("red", "white")))
cat("The following is the boxes matrix")
boxes
```

```{r}
p_red <- function(boxes) {
  Pr_boxes_balls <- boxes/rowSums(boxes)
  P_R_box <- matrix(Pr_boxes_balls[,1], ncol = 1)
  Pr_boxes <- matrix(c(.4,.1,.5), ncol = 3)
  p_red <- Pr_boxes %*% P_R_box
  return(p_red)
}

pred <- p_red(boxes)

cat("The probability of a red ball is P(R):", pred)
```

```{r}
p_box <- function(boxes) {
  Pr_boxes_balls <- boxes/rowSums(boxes)
  P_R_box <- matrix(Pr_boxes_balls[,1], ncol = 1)
  Pr_boxes <- matrix(c(.4,.1,.5), ncol = 3)
  P_box_R_list <- list()
  for(i in 1:3) {
    P_box_R <- (P_R_box[i]*Pr_boxes[i])/pred
    P_box_R_list[i] <- P_box_R
  }
  return(P_box_R_list)
}

cat("These are the probabilities of each box being picked given grabbing a red ball")
p_box(boxes)
cat("\n")
cat("The most likely box to be picked if we observed a red ball is c")
```

# Exercise (Bayes' theorem):

First thing to notice is that probabilities here dependent on Elvis and the other facts that we know. That is $P(B|I,Elvis) = 1$ because we know that Elvis was male. For sake of exposition, I dropped the $Elvis$ from the notation.

Our bayes theorem formula is:

$P(I|B) = \frac{P(B|I) P(I)}{P(B)} = \frac{P(B|I) P(I)}{P(B|I) P(I) + P(B|F) P(F)}$

where $I$ stands for identical twin, $B$ is for fraternal. Let us assume the following (no strict) assumption holds:

$P(B|F) = \frac{P(B,F)}{P(F)} = \frac{P(B) P(F)}{P(F)} = P(B)$

The latter simply assumes that the probability of Elvis having a fraternal brother is idependent of the probability of the brother being a boy.

Therefore, $P(I|B)$ is equal to:

```{r}
fraternal_prob <- 1/150
identical_prob <- 1/400
p_identifical_twin <- function() {
  identical_prob/(identical_prob + .5*fraternal_prob)
}
p_identifical_twin()
```









